---
title: "Data 606 Data Project Proposal"
author: "Stephen Haslett"
date: "10/14/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Research question 

Which college majors provide the highest return on investment?



### Cases 

There are 174 cases, each of which represents a college major.



### Data collection 

The original data was collected by the U.S. Census Bureau as part of the 2010-2012 American Community Survey (ACS). FiveThirtyEight (https://fivethirtyeight.com) used this data to construct a new dataset that acted as the factual foundation for their "The Economic Guide To Picking A College Major" (https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major) story. This data project will utilize FiveThirtyEight's version of the data.   



### Type of study 

This is an observational study examing college majors and their relation to employment and earnings.



### Data Source 

The dataset was created byFiveThirtyEight and can be found here - https://github.com/fivethirtyeight/data/blob/master/college-majors/all-ages.csv. The set was constructed with data from the "American Community Survey 2010-2012 Public Use Microdata Series" (https://www.census.gov/programs-surveys/acs/data/pums.html).



### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable in this analysis is the qualitative variable "Major_category".



### Independent Variable

The independent variables in this analysis are "Major_category", which is qualitative, and "Median", which is quantitative.




### Relevant summary statistics 

**Data Sample**

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

csv = "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv"
college_majors <- read.csv(file = csv, header = TRUE)

head(college_majors) %>% kable() %>% kable_styling(bootstrap_options = c('striped', 'bordered')) %>%
row_spec(1:6, color = "black")
```
