---
title: "DATA 606 Final Data Project"
author: "Stephen Haslett"
date: "11/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(stringr)
```

### Part 1 - Introduction


### Part 2 - Data

**To answer my research question, I will use 3 seperate datasets. One of the datasets was collected from Kaggle, and the other 2 were collected from FiveThirtyEight.**
**The datasets can be found at the following locations:**

1. Salary Increase by Major - https://www.kaggle.com/wsj/college-salaries/version/1#degrees-that-pay-back.csv

2. Employment Data for Recent Graduates - https://github.com/fivethirtyeight/data/blob/master/college-majors/recent-grads.csv

3. Employment Data for All Graduates - https://github.com/fivethirtyeight/data/blob/master/college-majors/all-ages.csv


```{r, message=FALSE, warning=FALSE}
category_csv <- 'https://raw.githubusercontent.com/stephen-haslett/data606/data606-final-project/final_data_project/degrees-that-pay-back.csv'
salary_by_major <- read.csv(url(category_csv), header = TRUE)
head(salary_by_major)
```


```{r, message=FALSE, warning=FALSE}
recent_grads_csv <- 'https://raw.githubusercontent.com/stephen-haslett/data606/data606-final-project/final_data_project/recent-grads.csv'
recent_grads_employment <- read.csv(url(recent_grads_csv), header = TRUE)
head(recent_grads_employment)
```


```{r, message=FALSE, warning=FALSE}
grad_employment_csv <- 'https://raw.githubusercontent.com/stephen-haslett/data606/data606-final-project/final_data_project/all-ages.csv'
all_grads_employment <- read.csv(url(grad_employment_csv), header = TRUE)
head(all_grads_employment)
```


### Part 3 - Exploratory data analysis


### Part 4 - Inference


### Part 5 - Conclusion